Distribution-oblivious Programming with Dish

### 2019-11-21 pldi/draft.tex

The key insight behind \sys is that the language of the Unix shell already encodes stream processing, providing most of the information required for distributing a computation.
\sys builds on this insight with a careful study of the scalability properties of shell primitives and its commands, paired with a novel rewriting scheme that identifies maximal pipeline sub-expressions that are candidates for scale-out.
We term the result \emph{distribution-oblivious programming}---a series of techniques that fully automates program distribution while provably maintaining correct (non-distributed) semantics.

Key benefits include the ability to scale legacy computations onto
\sys converts legacy shell pipelines into their distributed equivalents fully automatically, offering $100\times$ improvements in performance without a single line of additional code or annotation.
Most importantly, \sys provides an architectural lesson:
  the current design shows that large-scale efforts in the distributed- and operating-system literature to provide a \unix-like distributed equivalent could have been simplified by a thin (but sophisticated) rewriting shim like the one \sys provides.

>>>>

To answer this question, this paper presents \sys, a shell \todo{preprocessor} that transforms shell pipelines into their distributed equivalents enabling \emph{distribution-oblivious programming}.
The key insight behind \sys is that the language of the Unix shell already encodes stream processing, providing most of the information required to distribute a computation.
\sys builds on this insight with a careful study of the distributability properties of shell primitives and commands, a pipeline rewriting pass that identifies maximal sub-expressions as candidates for scale-out, and a late-bound command-prefixing scheme for deferring scheduling decisions to the distributed planner (\ie when runtime information is available).
A \sys-enabled $p_1$ will run \ttt{tr} and \ttt{uniq} in completely parallel streams, use a mostly-parallel tree of \ttt{sort}s, run \ttt{head} in one of the streams, and (trivially) merge streams in \ttt{out}.

The combination of automated transformations for program distribution with the ability to maintain correct non-distributed semantics results in several benefits.
% \sys converts legacy shell pipelines into their distributed equivalents fully automatically, offering $100\times$ improvements in performance without a single line of additional code or annotation.
First, shell users composing pipelines (or simply running legacy pipelines on massive datasets) can see scalability benefits without any manual effort---no need for \ttt{qsub}~\cite{gentzsch2001sun}, \textsc{SLURM}~\cite{yoo2003slurm}, calls to \textsc{GNU} \ttt{parallel}~\cite{Tange2011a}, or any manual rewriting~\cite{mapreduce:08, ciel:11, spark:12}.
Second, developers of new shell commands can use a concise a domain-specific language to express their distributability properties rather than using ad-hoc, command-specific flags such as {\tt -t},  {\tt NUM\_THREADS}, \ttt{-t}, \ttt{-p} \etc
Most importantly, \sys provides an architectural lesson for system designers---namely, that large-scale research efforts in the distributed- and operating-system literature to provide a \unix-like distributed equivalent~\cite{ousterhout1988sprite, mullender1990amoeba, pike1990plan9, barak1998mosix} would have been simplified by a thin (but sophisticated) rewriting shim like the one \sys provides.
% Nice critique on parallel: https://bugs.debian.org/cgi-bin/bugreport.cgi?bug=597050#75
